---
title: "Resources"
output:
  html_document:
    css: styles.css
    includes:
      after_body: footer.html
---


```{r tabfig, echo=FALSE}
source("R/tabfig.R")
```

## Books {#books}

```{r do-books, echo=FALSE}
width <- "160px"
tab(class="cellpadding", width="800px",
  tr(
  	tabfig("fox", "images/books/car-3e.jpg", 
  	       "https://us.sagepub.com/en-us/nam/an-r-companion-to-applied-regression/book246125", width=width),
  	tabtxt("Fox & Weisberg,",  a("An R Companion to Applied Regression",
  	                         href="https://us.sagepub.com/en-us/nam/an-r-companion-to-applied-regression/book246125"),
  	       ". A comprehensive introduction to linear models, regression diagnostics, etc.", br()
           # "Course notes at", aself("http://ccom.unh.edu/vislab/VisCourse/index.html")
  	       )
  	),

  tr(
  	tabfig("Wickham", "images/books/ggplot-book-2ndEd.jpg", 
  	       "https://www.springer.com/gp/book/9780387981413", width=width),
  	tabtxt("Hadley Wickham,",  a("ggplot2: Elegant Graphics for Data Analysis",
  	       href="https://www.springer.com/gp/book/9780387981413"),
  	       ". The printed version of the ggplot2 book. The 3rd edition is online at",
  	       a("https://ggplot2-book.org/", href= "https://ggplot2-book.org/")
           )
  	),
  tr(
  	tabfig("Chang", "images/books/RGraphicsCookbook.jpg", 
  	       "https://www.amazon.com/dp/1491978600", width=width),
  	tabtxt("Winston Chang,",  a("R Graphics Cookbook: Practical Recipes for Visualizing Data, 2nd Edition",
  	       href="https://www.amazon.com/dp/1491978600"),
  	       ". A problem-oriented book on R Graphics, with 'recipes' for common graphing problems. This edition is online at",
  	       a("https://r-graphics.org/", href= "https://r-graphics.org/")
           )
  	)

  )
```

### Other books on data visualization

For learning more about basic and advanced plotting using R:

* Kieran Healy, [Data Visualization - A practical introduction](https://socviz.co/). Contains an early draft of the book online.
* Antony Unwin, [Graphical Data Analysis with R](http://www.gradaanwr.net/). Web site contains all figures & R code.



<!-- ## Software {#software} -->

## R 

* Instructions to install R and R Studio for [Windows](info/installing-r-rstudio-windows.pdf) and [Mac](info/installing-r-rstudio-mac.pdf).
* An [interactive tutorial](https://learnr-examples.shinyapps.io/ex-setup-r/) that guides you through
the steps of installing R and R Studio.


## Other links

* Students in this course can book a [consulting appointment](http://www.appointmentquest.com/provider/2000199121) with the 
[Statistical Consulting Service](http://www.isryorku.ca/centres/scs/) for advice on data analysis for their research.  



