---
#############################################################
#                                                           #
# Click on "Run Document" in RStudio to run this worksheet. #
#                                                           #
#############################################################
title: "Visualizing Multivariate Data: Penguins"
author: "Michael Friendly"
output: learnr::tutorial
runtime: shiny_prerendered
---


```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE,
                      comment = "")
knitr::set_alias(w = "fig.width", h = "fig.height")

library(palmerpenguins)
library(learnr)
library(dplyr)
library(tidyr)

## Clean up variable names, get rid of NAs, etc.
# Remove units from variable names (for display purposes). Character variables should be factors.
peng <- penguins %>%
  rename(
    bill_length = bill_length_mm, 
    bill_depth = bill_depth_mm, 
    flipper_length = flipper_length_mm, 
    body_mass = body_mass_g
  ) %>%
  mutate(species = as.factor(species),
         island = as.factor(island),
         sex = as.factor(substr(sex,1,1))) %>%
  filter(!is.na(bill_depth),
         !is.na(sex))
```



## Introduction

The purpose of this exercise is to introduce you to some basic graphical displays
useful for multivariate data, particularly those associated with simple one-way
MANOVA designs.

We will be using the following R packages: 

* **dplyr**, **tidyr** for data wrangling,
* **ggplot2** for some plotting,
* **car** for better data analysis graphics,
* **heplots**, **candisc** for visualizing multivariate linear models

```{r library-calls, echo = TRUE}
# load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(car)
library(heplots)
library(candisc)
```

This exercise uses the [**learnr**](https://rstudio.github.io/learnr/) package.  

## Penguins data

The `Penguins` dataset, from the **palmerpenguins** package, contains "size" data on 
three species of penguins from three islands in Antartica. The size variables
are two measures of their bills (`bill_length`, and `bill_depth`), length of the
flipper, and body mass.

I have done some pre-processing to create a revised `peng` dataset, to: 

* make variable names simpler for display in plots (remove units), and
* deleting observations with missing data.

This code has already been run; it uses
uses functions from the **dplyr** and **tidyr** packages. You should be able
to read this and understand roughly what it is doing, but this topic is beyond the scope
of this course.

```{r eval=FALSE}
library(palmerpenguins)
peng <- penguins %>%
  rename(
    bill_length = bill_length_mm, 
    bill_depth = bill_depth_mm, 
    flipper_length = flipper_length_mm, 
    body_mass = body_mass_g
  ) %>%
  mutate(species = as.factor(species),
         island = as.factor(island),
         sex = as.factor(substr(sex,1,1))) %>%
  filter(!is.na(bill_depth),
         !is.na(sex))
```

We'll work with the resulting dataset, named `peng` here.

```{r peng, echo = TRUE}
car::some(peng, 8)
```

What R function can you use to find out the number of observations in this data set? 

```{r peng-q1, exercise=TRUE}
____(peng)
```

```{r peng-q1-solution}
nrow(peng)
```

Before going on, it is useful to see the number of observations for each species, 

```{r peng-species}
table(peng$species)
```

or broken down by another factor, like `island`. How would you extend this to show the
cross-classification of `species` and `island`?

```{r peng_sp_is, exercise=TRUE}
table(peng$species, peng$_____)
```

```{r peng_sp_is-solution}
table(peng$species, peng$island)
```



## Scatterplot matrix

To get an overview of the data, let's use `car::scatterplotMatrix` to show all pairwise plots of the
size variables.

```{r spm0, w=7, h=7}
scatterplotMatrix(~ bill_length + bill_depth + flipper_length + body_mass,
                  data=peng)
```

That's not very nice, because it conceals the differences among species.
Run this again, but conditioning the plots by `species`.

```{r spm1, exercise=TRUE}
scatterplotMatrix(~ bill_length + bill_depth + flipper_length + body_mass | _______ ,
                  data=peng)
```

```{r spm1-solution, w=7, h=7}
scatterplotMatrix(~ bill_length + bill_depth + flipper_length + body_mass | species ,
                  data=peng)
```

`scatterplotMatrix` has many options to control the details of what
is plotted. See: `help(scatterplotMatrix)`. I really like data ellipses
(`ellipse=TRUE`), and then it is less necessary to plot the data points 
(`plot.points=FALSE`).  Try this.

```{r spm2, exercise=TRUE}
scatterplotMatrix(~ bill_length + bill_depth + flipper_length + body_mass | species,
                  data = peng,
                  ellipse = ____,
                  plot.points = ____)
```

```{r spm2-solution}
scatterplotMatrix(~ bill_length + bill_depth + flipper_length + body_mass | species,
                  data = peng,
                  ellipse = TRUE,
                  plot.points = FALSE)
```

## Boxplots

Assume the goal is to determine whether/how the penguins differ in _size_ by `species`,
where the _size_ variables are `bill_length`, `bill_depth`, `flipper_length`, and `body_mass`.

For such data, comparative univariate displays (boxplots, violin plots, ...) of each response
by species provide a graphic overview.

For a single response variable, this is easy to do with `ggplot`,  using `geom_boxplot()` or
`geom_violin()`.  

```{r boxp1}
ggplot(peng, aes(x=species, y=body_mass, fill=species)) +
  geom_boxplot() +
  geom_jitter(position=position_jitter(0.1)) +
  theme(legend.direction = 'horizontal', legend.position = 'top') 
```

For more than one response variable, you can repeat this code, changing `y=body_mass` to use a different
variable. Try this for `bill_length`.

```{r boxp2, exercise=TRUE, w=10}
ggplot(peng, aes(x=species, y=_______, fill=species)) +
  geom_boxplot() +
  geom_jitter(position=position_jitter(0.1)) +
  theme(legend.direction = 'horizontal', legend.position = 'top') 
```


```{r boxp2-solution, w=10}
ggplot(peng, aes(x=species, y=bill_length, fill=species)) +
  geom_boxplot() +
  geom_jitter(position=position_jitter(0.1)) +
  theme(legend.direction = 'horizontal', legend.position = 'top') 
```

To see how the species compare on **all** size variables together, it is better to
compose such plots side-by-side.  This requires us to re-shape the data set from
"wide" format (responses in multiple colums) to "long" format (responses as multiple
rows).  

```{r peng-long}
peng_long <- peng %>% 
  tidyr::pivot_longer(bill_length:body_mass, 
                      names_to = "Measure", 
                      values_to="Size") 
```

Now, the boxplots for all responses can be described as a plot of `Size` against `species`, with
separate panels ("facets") for each `Measure`.

```{r boxp3, w=10}
ggplot(peng_long, aes(x=species, y=Size, fill=species)) +
  geom_boxplot() + 
  facet_wrap(. ~ Measure, scales="free_y", nrow=1)
```


## Univariate ANOVAs

Assume the goal is to determine whether/how the penguins differ in _size_ by `species`,
where the _size_ variables are `bill_length`, `bill_depth`, `flipper_length`, and `body_mass`.
We could do this for each variable separately.

Fit a univariate ANOVA model for `bill_length`, using `lm()`. Use the `car::Anova()` function to assess the effect
of `species`.

```{r lm_bill_len, exercise=TRUE}
lm1 <-lm(___ ~ ~~~, data=peng ) 
Anova(lm1)
```

```{r lm_bill_len-hint}
lm1 <-lm(bill_length ~ ~~~, data=peng ) 
Anova(lm1)
```

```{r lm_bill_len-solution}
lm1 <-lm(bill_length ~ species, data=peng ) 
Anova(lm1)
```

Can you do a similar analysis for `bill_depth` ?

```{r lm_bill_dep, exercise=TRUE}
lm2 <-lm(bill_depth ~ ~~~, data=peng ) 
Anova(lm2)
```

```{r lm_bill_dep-solution}
lm2 <-lm(bill_depth ~ species, data=peng ) 
Anova(lm2)
```


### Contrasts

There are $ng=3$ penguin species and the effect of `species` here, with $df=2$ degrees of freedom, can
be broken down into two separate tests of _contrasts_.

In R, this is accomplished by assigning relative weights to the groups in a matrix with ng rows and
df columns.

```{r}
contrasts(peng$species) <- matrix(c(1,-1, 0, 
                                   -1,-1,-2), nrow=3, ncol=2)
contrasts(peng$species)
```

Once this is done, separate tests of the contrasts can be obtained with `car::linearHypothesis()`
and plotted with `heplot()`.  



## MANOVA

To carry out a MANOVA, it is only necessary to include the separate response variables
as `cbind(Y1, Y2, ...)` on the left of the model formula.  Do this for the four _size_
variables to give a multivariate test.

```{r peng_mlm, exercise=TRUE}
peng.mlm <-lm(cbind(____, ____, ____, ____) ~ species, data=peng)
Anova(peng.mlm)
```

```{r peng_mlm-hint}
peng.mlm <-lm(cbind(bill_length, bill_depth, ____, ____) ~ species, data=peng)
Anova(peng.mlm)
```

```{r peng_mlm-solution}
peng.mlm <-lm(cbind(bill_length, bill_depth, flipper_length, body_mass) ~ species, data=peng)
Anova(peng.mlm)
```

## HE plots



```{r peng-he, w=6, h=6}
peng.mlm <-lm(cbind(bill_length, bill_depth, flipper_length, body_mass) ~ species, data=peng)
heplot(peng.mlm, fill=TRUE, fill.alpha=0.2)

```

